---
title: "trackingC.Rmd"
author: "Leeya Pressburger"
date: "`r Sys.Date()`"
output:
  bookdown::html_document2:
    fig_caption: yes
    number_sections: yes
    toc: true
    toc_depth: 2
    toc_float: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
```

### R setup

Load necessary packages, read in an ini file and initiate core, and set seed for reproducibility.

```{r intro, warning=FALSE, message=FALSE}
library(hector)
library(relaimpo)
library(dplyr)
library(tidyr)
library(ggplot2)
library(gridExtra)
library(ggExtra)
library(GGally)
library(bookdown)
library(data.table) 

theme_set(theme_light())

source("trackingC_common.R")

# Read in data files
runlist <- fread("./output_files/runlist.csv")
model_output <- fread("./output_files/model_output.csv")
trk_output <- fread("./output_files/trk_output.csv")
dest_output <- fread("./output_files/dest_output.csv")

# Add ssp information to the output data
# Will model_output ever NOT be sorted by run_number?
model_output$ssp <- runlist$ssp[model_output$run_number]
trk_output$ssp <- runlist$ssp[trk_output$run_number]
dest_output$ssp <- runlist$ssp[dest_output$run_number]

# If there are over 100 observations per time point and we want to plot
# individual runs, we use slice_sample() to randomly select 100 rows.

# Pull random run numbers from the runlist and filter output by random runs
# We do replace=TRUE in case there are fewer than 100 runs available
max_display <- slice_sample(runlist, n = 100, replace = TRUE)

```

### Results

The following sections contain tentative figures in the order they would appear
in a manuscript. 

# Primary figures

## Establishing figures 

### Hector atmospheric CO~2~ vs observed data

``` {r ca-comp, fig.cap = " Historical CO2 data from Hector runs and NOAA observations."}
# # Comparing outputs to NOAA CO2 data from Mauna Loa observatory
# # https://gml.noaa.gov/ccgg/trends/data.html
# noaa <- read.csv("./data_files/noaa_co2_data.csv")
# noaa_co2 <- noaa %>% 
#   mutate(source = "NOAA") %>%
#   filter(year < 2015)

## First, we need to read in the CMIP6 data and organize
# Read in CMIP CO2 data, compute summary statistics
cmip_co2 <- read.csv("./data_files/CMIP6_annual_co2.csv")

# Historical data
cmip_co2_hist <- cmip_co2 %>%
  filter(year %in% 1850:2014,
         experiment == "historical") %>%
  mutate(value = value * 1e6,
         units = "ppm") %>%
  group_by(year) %>%
  summarize(sdev = sd(value), 
            minimum = min(value),
            maximum = max(value),
            med = median(value),
            source = "CMIP6",
            .groups = "drop")

# Identify unrealistic runs - for each run, if more than 50% of values are
# outside of the min/max of the CMIP data, the run is nixed
cmip_bounds_h <- cmip_co2_hist %>% 
  filter(year %in% 1850:2014) %>%
  select(year, minimum, maximum) %>%
  ungroup()

# Get CO2 data from Hector, add "fail" column if values are outside the
# min or max of CMIP
carbon_h <- model_output %>%
  filter(variable == "CO2_concentration",
         year >= 1850 & year <= 2014) %>%
  select(run_number, year, value) %>%
  left_join(cmip_bounds_h, by = "year") %>%
  mutate(fail = value < minimum | value > maximum)

# Summarize
run_fail_summary_h <- carbon_h %>%
  group_by(run_number) %>%
  summarize(fraction_fail = sum(fail) / n())

# Filter out 'unrealistic' runs (identified above)
model_output_filtered_1 <- model_output %>%
  left_join(run_fail_summary_h, by = "run_number") %>% 
  filter(fraction_fail < 0.33) %>% 
  select(-fraction_fail)

rm(model_output)

# Note that the tracking output needs to be filtered as well
trk_output_filtered_1 <- trk_output %>%
  left_join(run_fail_summary_h, by = "run_number") %>% 
  filter(fraction_fail < 0.33) %>% 
  select(-fraction_fail)

rm(trk_output)

## Plot
# Read in Hector data, compute summary statistics 
hector_co2 <- model_output_filtered_1 %>%
  filter(variable == "CO2_concentration",
         year %in% 1850:2014,
         ssp == MAIN_SCENARIO) %>%
  select(run_number, year, value) %>%
  mutate(units = "ppm") %>%
  group_by(year) %>%
  summarize(sdev = sd(value), 
            minimum = min(value),
            maximum = max(value),
            med = median(value),
            source = "Hector",
            .groups = "drop")

plot_data_co2 <- bind_rows(cmip_co2_hist, hector_co2)

# Plot
co2_plot <- plot_data_co2 %>% 
  filter(year < 2015) %>%
  ggplot(aes(year, fill = source, color = source)) +
  geom_ribbon(aes(ymin = minimum, ymax = maximum), alpha = 0.15, color = NA) +
  geom_ribbon(aes(ymin = med - sdev, ymax = med + sdev), alpha = 0.5, color = NA) +
  geom_line(aes(y = med), size = 2) +
 # geom_line(data = noaa_co2, aes(year, mean, color = source), size = 1) +
  scale_color_viridis_d(begin = 0.5) +
  scale_fill_viridis_d(begin = 0.5) +
  labs(x = "Year",
       y = "CO2 (ppm)",
       color = "Source",
       fill = "Source") +
  scale_x_continuous(expand = c(0, 0)) ; co2_plot
```


### Hector temperature vs. CMIP6 

``` {r filtered, warning = FALSE, message = FALSE, fig.cap = "Global temperature anomaly from CMIP6 and filtered Hector runs. Lines show median; shaded areas show ±1 s.d. of median (darker) and minimum and maximum of ensemble (lighter)."}
# Read in CMIP6 comparison data
cmip_raw <- read.csv("./data_files/CMIP6_annual_tas_global.csv")

# Isolate historical data
cmip_tgav_hist <- cmip_raw %>% 
  filter(experiment == "historical",
         variable == "Tgav",
         year <= 2100) %>%
  group_by(year) %>%
  summarize(sdev = sd(value),
            minimum = min(value),
            maximum = max(value),
            med = median(value)) %>%
  mutate(source = "CMIP6") 

# Isolate desired scenarios
cmip_tgav_sce <- list()
for(n in names(SSP_files)){
  cmip_tgav_sce[[n]] <- cmip_raw %>% 
    filter(experiment == n)
}  

cmip_sce <- bind_rows(cmip_tgav_sce)
rm(cmip_raw)

# Filter for Tgav and calculate summary metrics
cmip_t_scenarios <- cmip_sce %>%
  rename(ssp = experiment) %>%
  filter(variable == "Tgav",
         year <= 2100) %>%
  group_by(year, ssp) %>%
  summarize(sdev = sd(value),
            minimum = min(value),
            maximum = max(value),
            med = median(value)) %>%
  mutate(source = "CMIP6") %>%
  arrange(ssp, year)

# Need to append the historical values to each SSP
# Add SSP column to historical data, repeat for all scenarios
cmip_hist_repeat_t <- list()
for(n in names(SSP_files)){
  cmip_hist_repeat_t[[n]] <- cmip_tgav_hist %>% 
    mutate(ssp = n)
}
cmip_hist_repeat <- bind_rows(cmip_hist_repeat_t)

cmip_t_final <- bind_rows(cmip_hist_repeat, cmip_t_scenarios) %>% arrange(ssp, year)

### Now that we have all the data we'll need in one place, we want to filter out run numbers twice.
# First, we just want to look at the historical data. Then we'll look at each SSP

# Identify unrealistic runs - for each run, if more than 50% of values are
# outside of the min/max of the CMIP data, the run is nixed
cmip_bounds_t <- cmip_t_final %>% 
  group_by(ssp) %>%
  select(year, minimum, maximum) %>%
  ungroup()

# Get temperature data from Hector, add "fail" column if values are outside the
# min or max of CMIP
tgav_h <- model_output_filtered_1 %>%
  filter(variable == "global_tas",
         year >= 1850 & year <= 2014) %>%
  select(run_number, ssp, year, value) %>%
  left_join(cmip_bounds_t, by = c("year", "ssp")) %>%
  mutate(fail = value < minimum | value > maximum)

# Summarize
run_fail_summary_tgav <- tgav_h %>%
  group_by(ssp, run_number) %>%
  summarize(fraction_fail = sum(fail) / n())

# Filter out 'unrealistic' runs (identified above)
model_output_filtered_2 <- model_output_filtered_1 %>%
  left_join(run_fail_summary_tgav, by = c("ssp", "run_number")) %>% 
  filter(fraction_fail < 0.33) %>%
  select(-fraction_fail)

rm(model_output_filtered_1)

# Note that the tracking output needs to be filtered as well
trk_output_filtered_2 <- trk_output_filtered_1 %>%
  left_join(run_fail_summary_tgav, by = c("ssp", "run_number")) %>% 
  filter(fraction_fail < 0.33) %>%
  select(-fraction_fail)

rm(trk_output_filtered_1)

# Repeat for SSPs - future projections
# Identify unrealistic runs - for each run, if more than 50% of values are
# outside of the min/max of the CMIP data, the run is nixed
# Get temperature data from Hector, add "fail" column if values are outside the
# min or max of CMIP
tgav_h_ssp <- model_output_filtered_2 %>%
  filter(variable == "global_tas",
         year >= 2015 & year <= 2100) %>%
  select(run_number, ssp, year, value) %>%
  left_join(cmip_bounds_t, by = c("year", "ssp")) %>%
  mutate(fail = value < minimum | value > maximum)

# Summarize
run_fail_summary_tgav_ssp <- tgav_h_ssp %>%
  group_by(ssp, run_number) %>%
  summarize(fraction_fail = sum(fail) / n())

# Filter out 'unrealistic' runs (identified above)
model_output_final <- model_output_filtered_2 %>%
  left_join(run_fail_summary_tgav_ssp, by = c("ssp", "run_number")) %>% 
  filter(fraction_fail < 0.5)

rm(model_output_filtered_2)

# Note that the tracking output needs to be filtered as well
trk_output_final <- trk_output_filtered_2 %>%
  left_join(run_fail_summary_tgav_ssp, by = c("ssp", "run_number")) %>% 
  filter(fraction_fail < 0.5)

rm(trk_output_filtered_2)

# Create facet labels
ssp_labels <- c("SSP126", "SSP245", "SSP370", "SSP460", "SSP585")
names(ssp_labels) <- c(names(SSP_runs))

# Plot temperature spreads for each scenario, Hector vs CMIP
temp_plot <- model_output_final %>%
  filter(variable == "global_tas",
         year >= 1850 & year <= 2100) %>%
  # compute median, min, max, etc. for all remaining runs
  group_by(year, ssp) %>%
  summarize(sdev = sd(value),
            minimum = min(value),
            maximum = max(value),
            med = median(value), .groups = "drop") %>% 
  # combine with CMIP6 data
  mutate(source = "Hector") %>%
  bind_rows(cmip_t_final) %>%
  # ...and plot
  ggplot(aes(year, fill = source, color = source)) +
  geom_ribbon(aes(ymin = minimum, ymax = maximum), alpha = 0.15, color = NA) +
  geom_ribbon(aes(ymin = med - sdev, ymax = med + sdev), alpha = 0.5, color = NA) +
  geom_line(aes(y = med), size = 1) +
  scale_color_viridis_d(begin = 0.5) +
  scale_fill_viridis_d(begin = 0.5) +
  facet_wrap(~ssp, scales = "free",
             labeller = labeller(ssp = ssp_labels),
             ncol = 3) +
  #theme(legend.title = element_blank()) +
  labs(x = "Year",
       y = expression(degree*C),
       color = "Source",
       fill = "Source") +
  scale_x_continuous(expand = c(0, 0)) ; temp_plot
```

## Hector temperature vs CMIP6 diagnostic

``` {r s-cmip, message = FALSE, warning = FALSE, fig.cap = " Visualization of sample model runs. The black dots represent the CMIP6 minimums and maximums for global temperature anomaly, and the runs that pass must fall within these bounds more than 50% of the time."}
# Visualize the run fail summary calculated above
# Make a table of failure rates
run_fail_summary_h %>% 
  group_by(fraction_fail > 0.33) %>% 
  summarise(n()) %>% 
  knitr::kable()

run_fail_summary_tgav %>% 
  group_by(fraction_fail > 0.33) %>% 
  summarise(n()) %>% 
  knitr::kable()

run_fail_summary_tgav_ssp %>% 
  group_by(fraction_fail > 0.5) %>% 
  summarise(n()) %>% 
  knitr::kable()

# ...and a figure 
hector_temp_runs <- tgav_h_ssp %>% 
  group_by(ssp) %>%
  filter(year == min(year)) %>%
  slice_sample(n = 100, replace = TRUE) %>% # this is a debugging figure; plot just 100 runs for clarity
  arrange(ssp, run_number) %>%
  summarize(run_number)

# Filter out data for just the 100 runs per scenario identified above
hector_temp_slice <- left_join(hector_temp_runs, tgav_h_ssp, by = c("run_number", "ssp"))
  
# Plot
left_join(hector_temp_slice, run_fail_summary_tgav_ssp, by = c("run_number", "ssp")) %>%
  ggplot(aes(year, value, color = fraction_fail, group = run_number)) + 
  geom_line() + 
  geom_point(aes(y = minimum), color = "black") + 
  geom_point(aes(y = maximum), color = "black") +
  facet_wrap(~ssp,
             labeller = labeller(ssp = ssp_labels),
             ncol = 3) +
  scale_color_viridis_c() +
  labs(x = "Year",
       y = expression(degree*C),
       color = "Fraction of failed runs")

```

``` {r cleanup-cmip, echo = FALSE}
rm(list = c("cmip_co2", "cmip_co2_hist", "cmip_bounds_h", "cmip_bounds_t", "cmip_hist_repeat",
     "cmip_hist_repeat_t", "cmip_sce", "cmip_t_final", "cmip_t_scenarios",
     "cmip_tgav_hist", "cmip_tgav_sce"))
```

## Sources of atmospheric CO~2~

### Sources by pool

``` {r atmosphere-by-source, message = FALSE, fig.cap = paste0(" Fraction of CO2 in the atmosphere by source pool in ", MAIN_SCENARIO, ". Lines show median; shaded areas show ±1 s.d. of median (darker) and minimum and maximum of ensemble (lighter).")}
# Isolate the atmosphere pool to better understand its composition 
atmosphere_pool <- trk_output_final %>%
  filter(pool_name == "atmos_co2")

# Create facet labels
pool_labels <- c("Atmosphere", "Deep ocean", "Detritus", "Anthropogenic emissions",
                               "High-lat ocean", "Intermed. ocean",
                               "Low-lat ocean", "Soil", "Vegetation")
names(pool_labels) <- c("atmos_co2", "deep", "detritus_c", "earth_c", "HL", "intermediate",
                               "LL", "soil_c", "veg_c")

# This plot looks at source fraction by source pool in the atmosphere and the median run.
atm_summary <- atmosphere_pool %>%
  filter(ssp == MAIN_SCENARIO) %>%
  group_by(year, source_name) %>%
  summarize(sf_sd = sd(source_fraction), 
            sf_min = min(source_fraction),
            sf_max = max(source_fraction),
            sf_median = median(source_fraction))

atm_summary$source_name <- factor(atm_summary$source_name, levels = rev(c("HL", "LL", "detritus_c",
                                                                          "intermediate", "deep", "veg_c",
                                                                          "soil_c", "earth_c", "atmos_co2")))
atm_plot <- atm_summary %>%
  filter(year > 1800) %>%
  ggplot(aes(year)) +
  geom_ribbon(aes(ymin = sf_min, ymax = sf_max), alpha = 0.2) +
  geom_ribbon(aes(ymin = sf_median - sf_sd, ymax = sf_median + sf_sd), alpha = 0.2) +
  geom_line(aes(y = sf_median)) +
  facet_wrap(~source_name, scales = "free",
             labeller = labeller(source_name = pool_labels)) +
  labs(x = "Year",
       y = "Source fraction of CO2 in atmosphere") +
  scale_x_continuous(expand = c(0, 0)) ; atm_plot
 

percent_summary <- atmosphere_pool %>%
  filter(year %in% c(2010, 2050, 2100, 2300),
         source_name == "earth_c") %>%
  group_by(ssp, year) %>%
  summarize(avg = mean(source_fraction),
            med = median(source_fraction))
  
# atmosphere_pool %>%
# filter(ssp == MAIN_SCENARIO,
#        run_number %in% max_display$run_number) %>%
# ggplot(aes(year, source_fraction, 
#              color = as.factor(run_number), group = run_number)) +
#   geom_line(size = 0.5, show.legend = FALSE) +
#   facet_wrap(~source_name, scales = "free",
#              labeller = labeller(source_name = pool_labels)) +
#   labs(x = "Year",
#        y = "Source fraction") +
#   scale_color_grey(start = 0.7, end = 0.7) +
#   stat_summary(color = "black",
#                fun = median,
#                geom = "line",
#                group = "run_number",
#                size = 0.7)
```

### Anthropogenic emissions in the atmosphere across scenarios

``` {r earthc-in-atmos, fig.cap = paste0(" Fraction of atmospheric CO2 from anthropogenic emissions over time across scenarios. Line shows median; shaded areas show ±1 s.d. of median (darker) and minimum and maximum of ensemble (lighter).")}
# This chunk looks at the atmosphere pool with just one source (earth_c) 
# and plots source fraction over time with a confidence interval

ff <- atmosphere_pool %>% 
  filter(source_name == "earth_c") %>%
  group_by(ssp, year) %>%
  mutate(sf_sd = sd(source_fraction), 
            sf_min = min(source_fraction),
            sf_max = max(source_fraction),
            sf_median = median(source_fraction))

atm_fraction_plot <- ggplot(ff, aes(year,  fill = ssp)) +
  geom_line(aes(y = sf_median, color = ssp)) +
  #geom_ribbon(aes(ymin = sf_min, ymax = sf_max), alpha = 0.2) +
  geom_ribbon(aes(ymin = sf_median - sf_sd, ymax = sf_median + sf_sd), alpha = 0.2) +
  scale_color_viridis_d(breaks = names(ssp_labels),
                        labels = ssp_labels) +
  scale_fill_viridis_d(breaks = names(ssp_labels),
                        labels = ssp_labels) +
  labs(x = "Year",
       y = "Fraction of atmospheric CO2",
       color = "Scenario",
       fill = "Scenario") +
  scale_x_continuous(expand = c(0, 0)) ; atm_fraction_plot
```

``` {r cleanup-source}
rm(list = c("atmosphere_pool", "ff", "atm_summary", "atm_plot", "percent_summary"))
```

## Destination of anthropogenic emissions

### Destination fraction

```{r destination, warning = FALSE, fig.cap = paste0(" Final destination pools of anthropogenic emissions averaged over ", SSP_runs[MAIN_SCENARIO], " runs in ", MAIN_SCENARIO, ".")}
# Where does CO2 end up?

# Just looking at 5 random runs across scenarios
# dest_runs <- ffi_destinations %>%
#   group_by(ssp) %>%
#   filter(year == min(year)) %>%
#   slice_sample(n = 5) %>%
#   summarize(run_number)

# dest_plot <- filter(ffi_destinations, 
#                     run_number %in% dest_runs$run_number) 

# Ordered pool labels
o_pool_labels <- c("Atmosphere", "Soil", "Vegetation",  "Detritus",
                                  "High-lat ocean", "Low-lat ocean",
                                  "Intermed. ocean", "Deep ocean")
names(o_pool_labels) <- c("atmos_co2", "soil_c", "veg_c",  "detritus_c",
                                  "HL", "LL", "intermediate", "deep")

# Just looking at the mean of all runs in SSP245
dest_plot <- dest_output %>%
 # filter(ssp == MAIN_SCENARIO) %>%
  group_by(ssp, year, pool_name) %>%
  mutate(dest_avg = mean(destination_fraction),
         pool_name = factor(pool_name, levels = rev(c("deep", "intermediate",
                                                      "LL", "HL", "detritus_c", 
                                                      "veg_c", "soil_c", "atmos_co2")))) %>%
  filter(run_number == min(run_number),
         year > 1800) %>%
  ggplot(aes(year, dest_avg, fill = pool_name)) +
  geom_area() +
  scale_fill_viridis_d(breaks = names(o_pool_labels),
                       labels = o_pool_labels) +
  labs(x = "Year",
       y = "Fraction of anthropogenic emissions",
       fill = "Destination") +
  scale_x_continuous(expand = c(0, 0)) +
  # These lines are to look at multiple scenarios
  facet_wrap(~ssp,
             labeller = labeller(ssp = ssp_labels)) +
  theme(#strip.background = element_blank(),
    # strip.text.x = element_blank(),
    axis.text.x = element_text(angle = 90)) ; dest_plot

```

### Does time affect destination?

```{r tracking-multiple-years, warning = FALSE, fig.cap = paste0(" Final destination pools of anthropogenic emissions in ", MAIN_SCENARIO, " for ten 50-year periods between 1750 and 2250. Note that turning on carbon tracking at the beginning of each period does not have a dramatic effect on the destination of emissions.")}
# Prepare data: reset the core, set start dates every 50 years
tcore <- newcore(SSP_files[[MAIN_SCENARIO]])
dates <- seq(from = TRACKING_START, to = 2250, 50)
tracking_output <- list()

# Function to set a tracking date, run the core for a 50 year period, and retrieve tracking data
set_tracking <- function(c, date){
  setvar(c, NA, TRACKING_DATE(), date, "(unitless)")
  reset(c)
  run(c, (date + 50))
  tdata <- get_tracking_data(c)
  tdata$ty <- date
  tdata$ssp <- MAIN_SCENARIO
  tdata
}

# Loop through above function for desired dates
for(d in dates){
  tracking_output[[d]] <- set_tracking(tcore, d)
}

all_years_tracking <- bind_rows(tracking_output)

# Plot the fraction of the atmosphere that is sourced from earth_c
# all_years_tracking %>%
#   filter(source_name == "earth_c",
#          pool_name == "atmos_co2") %>%
#   ggplot(aes(year, source_fraction, color = ty, group = ty)) +
#   geom_line() +
#   labs(x = "Year",
#        y = "Source fraction") +
#   scale_color_viridis_c()

  
# Plot the destination of earth_c emissions
all_years_tracking %>%
  filter(source_name == "earth_c", pool_name != "earth_c") %>%
  mutate(source_quantity = pool_value * source_fraction) %>%
  group_by(year) %>%
  mutate(destination_fraction = source_quantity / sum(source_quantity)) %>% 
  ungroup() %>%
  mutate(pool_name = factor(pool_name, levels = rev(c("deep", "intermediate",
                                                      "LL", "HL", "detritus_c",
                                                      "veg_c", "soil_c", "atmos_co2")))) %>%
  filter(ty != "2250") %>%
  ggplot(aes(year, destination_fraction, fill = pool_name), group = ty) +
  geom_area() +
  facet_wrap(~ty, 
             scales = "free_x",
             nrow = 2) +
  labs(x = "Year",
       y = "Fraction of anthropogenic emissions",
       fill = "Destination") +
  scale_fill_viridis_d(breaks = names(o_pool_labels),
                       labels = o_pool_labels) +
  theme(axis.text.x = element_text(angle = 90))
```

### What controls how much anthropogenic CO2 ends up in the atmosphere?

```{r relative-importance, message=FALSE, echo=FALSE, warning=FALSE, fig.cap = " Relative importance of parameters over time in controlling the atmosphere as a destination of anthropogenic emissions. Relative importance is the R^2 contribution averaged over orderings among regressors; here, we normalize the values so that they sum to one."}
# Process data
dest <- dest_output %>% 
  filter(run_number %in% trk_output_final$run_number) %>%
  left_join(runlist, by = c("run_number", "ssp")) %>%
  as.data.frame()

# Filter to just atmosphere pool, and remove unneeded columns
dest_e2a <- dest %>% 
  filter(pool_name == "atmos_co2")
  
dest_e2a_minimal <- dest_e2a[c("year", "destination_fraction", "ssp", names(name_vector))]
rm(dest_e2a)

# Function to calculate relative importance of parameters in
# explaining how much destination_fraction varies for a given year of data. 
# This uses relaimpo::calc.relimp()
# See https://www.r-bloggers.com/2012/08/the-relative-importance-of-predictors-let-the-games-begin/
# for a useful primer, along with the calc.reimp help page
calc_relimp <- function(x) {
  yr <- x$year[1] # record year...
  x$year <- NULL # and drop year column
  sce <- x$ssp[1]
  x$ssp <- NULL
  # Fit model. The formula notation means "destination fraction as a
  # function of all other terms" (which is why we dropped year)
  m <- lm(destination_fraction ~ ., data = x)
  # Calculate relative importance metrics and extract 'lmg' results
  # 'lmg' is the R^2 contribution averaged over orderings among regressors;
  # should sum to one because we're using relative importances (rela=TRUE)
  relimp <- try(calc.relimp(m, type = "lmg", rela = TRUE)@lmg)
  if(class(relimp) == "try-error") {
    message(yr, " calc.relimp error")
    return(NULL)
  }
  # Return a data frame: year, parameter, relative importance
  tibble(ssp = sce, year = yr, param = names(relimp), value = relimp)
  
}

# Run calc_relimp() above on each year's data
dest_split <- split(dest_e2a_minimal, list(dest_e2a_minimal$year, dest_e2a_minimal$ssp))
relimp_out <- lapply(dest_split, FUN = calc_relimp) %>% 
  bind_rows()
    
# ...and plot!
importance_plot <- relimp_out %>%
  mutate(param = factor(param, levels = rev(c("NPP_FLUX0", "Q10_RH", "BETA", "ECS", 
                                       "DIFFUSIVITY", "AERO_SCALE",
                                       "LUC_SCALE")))) %>%
  filter(year > 1999) %>%
  ggplot(aes(year, value, fill = param)) + 
  geom_area() +
  coord_cartesian(expand = FALSE) +
  scale_fill_viridis_d(direction = -1,
                       begin = 0.1,
                       breaks = c("LUC_SCALE", "AERO_SCALE", "DIFFUSIVITY",
                                  "ECS", "BETA", "Q10_RH","NPP_FLUX0"),
                       labels = c("LUC emissions", "Aerosol forcing",
                                  "Ocean diffusivity", "ECS",
                                  "CO2 fertilization", "Q10",
                                  "Pre-industrial NPP")) +
  facet_wrap(~ssp,
             labeller = labeller(ssp = ssp_labels),
             ncol = 3) +
  labs(x = "Year",
       y = "Relative importance for anthropogenic CO2 in the atmosphere",
       fill = "Parameter") +
  theme(axis.text.x = element_text(angle = 90)) +
  scale_x_continuous(expand = c(0, 0)) ; importance_plot

```

``` {r cleanup-destination, echo = FALSE}
rm(list = c("dest_plot", "all_years_tracking", "dest", "dest_e2a_minimal", "dest_output",
     "dest_split", "importance_plot"))
```

## Airborne fraction

### Airborne fraction and "fraction of emissions remaining in atmosphere"

```{r airborne, warning = FALSE, fig.cap = " Airborne fraction over time, by scenario. The yellow line and band shows the actual amount of anthropogenic emissions remaining in the atmosphere from the model's carbon-tracking mechanism, whereas the blue line shows the actual airborne fraction."}
# Prep the data
atm_earth_diffs <- trk_output_final %>%
  # Isolate atmosphere and earth total pool values
  filter(pool_name %in% c("atmos_co2", "earth_c")) %>% 
  group_by(run_number, year, pool_name, ssp) %>% 
  summarise(pool_value = mean(pool_value), .groups = "drop") %>% 
  arrange(year) %>%
  # Put atmosphere and earth in separate columns and compute diffs
  pivot_wider(names_from = "pool_name", values_from = "pool_value") %>% 
  group_by(run_number, ssp) %>%
  mutate(atm_diff = c(NA, diff(atmos_co2)), 
         earth_diff = c(NA, diff(earth_c)))

# Compute AF (classical definition of delta CO2atm / emissions) 
# using various window sizes
af_results <- list()
for(groupsize in seq(1, 40, length.out = 10)) {
  ngroups = length(unique(atm_earth_diffs$year)) / groupsize
  
  atm_earth_diffs %>% 
    group_by(run_number, ssp) %>% 
    mutate(group = cut(year, breaks = ngroups)) %>% 
    # compute differences over the group of years
    group_by(run_number, group, ssp) %>%
    summarise(year = max(year), 
              atm_diff = sum(atm_diff), 
              earth_diff = sum(earth_diff), 
              n = n(),
              af = atm_diff / -earth_diff,
              .groups = "drop") ->
    af_results[[as.character(groupsize)]]
}

af_results <- bind_rows(af_results, .id = "window_size") %>% 
  mutate(window_size = as.integer(window_size))

# Compute cumulative emissions over time
# We use the change in earth_c to compute emissions; note this will not work
# if ever using a scenario with fossil fuel sequestration (atmosphere -> earth)

# This function computes annual cumulative earth_c emissions for a given tracking dataset
min_runs <- trk_output_final %>%
  group_by(ssp) %>%
  summarize(mrun = min(run_number))

calc_emissions <- function(data){
  data %>% 
    filter(pool_name == "earth_c",
         source_name == "earth_c",
         run_number %in% min_runs$mrun) %>%
    arrange(year) %>%
    group_by(ssp) %>%
    # the NA at the end here is so that cumulative emissions 'line up' correctly
    # with the source_quantity (earth-origin atmosphere C) calculated below
    mutate(emissions = -c(diff(pool_value), NA),
           # relative to core$trackingDate
           cumulative_emissions = cumsum(emissions)) %>%
    select(year, ssp, cumulative_emissions) %>%
    arrange(ssp, year)
}

# This function takes a given tracking dataset and its corresponding cumulative 
# emissions dataset as found above to calculate what fraction of the cumulative
# emissions come from earth_c emisions that are now in the atmosphere
calc_emissions_fraction <- function(data, emissions_data){
  data %>%
    filter(pool_name == "atmos_co2", 
           source_name == "earth_c") %>%
    # How much earth_c is in atmos_co2? 
    mutate(source_quantity = pool_value * source_fraction) %>%
    left_join(emissions_data, by = c("year", "ssp")) %>%
    select(run_number, ssp, year, source_quantity, cumulative_emissions) %>%
    mutate(c_emissions_fraction = source_quantity / cumulative_emissions) %>% 
    group_by(year, ssp) %>%
    mutate(sdev = sd(c_emissions_fraction),
           med = median(c_emissions_fraction),
           minimum = min(c_emissions_fraction),
           maximum = max(c_emissions_fraction),
           def = "Carbon tracking")
}

# Compute cumulative emissions over time
emissions <- trk_output_final %>% calc_emissions() 

# Compute the fraction of fossil fuel emissions residing in the atmosphere
emissions_fraction <- calc_emissions_fraction(trk_output_final, emissions)

# Classical definition of AF - find summary stats
classic_af <- af_results %>%
  group_by(group, year, ssp) %>%
  mutate(sdev = sd(af, na.rm = TRUE),
         minimum = min(af),
         maximum = max(af),
         med = median(af),
         def = "Airborne fraction") %>% 
  filter(window_size == 1,
         run_number %in% emissions_fraction$run_number)

# Isolate relevant parameters and plot
af_plot_data <- bind_rows(emissions_fraction, classic_af) %>%
  select(run_number, year, ssp, minimum, maximum, med, sdev, def) %>%
  filter(year %in% 1910:2200) # past 2200, data gets weird

af_plot <- ggplot(af_plot_data, aes(x = year, fill = def)) +
  geom_ribbon(aes(ymin = minimum, ymax = maximum), 
              alpha = 0.2) +
  geom_ribbon(aes(ymin = med - sdev, ymax = med + sdev), 
              alpha = 0.4) +
  geom_line(aes(y = med, color = def)) +
  scale_color_viridis_d(begin = 0.5) +
  scale_fill_viridis_d(begin = 0.5) +
  facet_wrap(~ssp, scales = "free_y", 
             labeller = labeller(ssp = ssp_labels),
             ncol = 3) +
  labs(x = "Year",
      y = "Airborne fraction",
      color = "Computation",
      fill = "Computation") +
  theme(axis.text.x = element_text(angle = 90)) +
  scale_x_continuous(expand = c(0, 0)) ; af_plot


average_af <- classic_af %>%
  ungroup() %>%
  filter(year %in% 1960:2020) %>%
  summarize(avg = mean(af),
            .groups = "drop")

```

### Airborne fraction trend

``` {r af-trend, message = FALSE, fig.cap = " The spread of airborne fraction decadal trend values from 1960-2020 across all runs and scenarios. See the table below for summary statistics."}
# Calculate airborne fraction trend from 1960-2020 by decade, as in van Marle et al. (2022)
# Extract results from above
af <- af_results %>%
  group_by(group, year, ssp) %>%
  mutate(def = "Airborne fraction") %>% 
  filter(window_size == 1,
         run_number %in% emissions_fraction$run_number,
         year %in% 1910:2200) %>%
  select(run_number, year, ssp, af, def)

# Calculate AF trend over time from 1960-2020
af_trend <- af %>%
  filter(year %in% 1960:2020) %>%
  group_by(run_number, def) %>%
  do(mod = lm(af ~ year, data = .)) %>%
  summarize(run_number = run_number,
            def = def,
            broom::tidy(mod)) %>%
  # Extract only the slope, not the intercept term
  filter(term == "year") %>%
  # Multiply annual trend by 10 for decadal average
  mutate(decade = estimate * 10,
         error = std.error * 10)

# Organize data, make table and graph
a <- af_trend %>%
  arrange(def) %>%
  select(run_number, 
         computation = def, 
         af_trend = decade, error, p.value) 

af_trend_plot <- a %>% 
  ggplot(aes(af_trend)) + 
  geom_histogram(bins = 40) +
  labs(x = expression(Airborne~fraction~trend~(decade^-1)),
       y = "Count") +
  scale_x_continuous(expand = c(0, 0)) ; af_trend_plot

# Summary stats
a_stats <- a %>%
  summarize(sdev = sd(af_trend),
         minimum = min(af_trend),
         maximum = max(af_trend),
         mean = mean(af_trend),
         median = median(af_trend))

a_stats %>% knitr::kable(digits = 4)

af_negative <- round(sum(a$af_trend < 0) / nrow(a) * 100, 1)

# LUC comparison
luc <- model_output_final %>%
  filter(year %in% 1960:2020,
         variable == "luc_emissions")

# Compute average LUC emissions by scenario  
luc_avg <- luc %>%
  group_by(ssp) %>%
  summarize(avg = mean(value))

# Compute LUC emissions trend, join with luc_avg
luc_trend <- luc %>%
  group_by(ssp) %>%
  do(mod = lm(value ~ year, data = .)) %>%
  summarize(ssp = ssp,
            broom::tidy(mod)) %>%
  # Extract only the slope, not the intercept term
  filter(term == "year") %>%
  left_join(luc_avg, by = "ssp")

# Organize data, make table
# luc_trend %>%
#   select(ssp, 
#          luc_trend = estimate, 
#          error = std.error,
#          avg) %>%
#   knitr::kable()

```

`r af_negative`% of the runs had a negative AF trend.

``` {r cleanup-af, echo = FALSE}
rm(list = c("atm_earth_diffs", "af_results", "min_runs", "emissions", 
            "emissions_fraction", "classic_af", "af_plot_data", "average_af", 
            "af", "af_trend", "af_trend_plot", "a_stats", "af_negative",
            "luc", "luc_avg", "luc_trend"))

```

# Supplementary figures

## Realized parameter PDFs

``` {r s-params-PDFs, fig.cap = paste0(" Probability densities of each parameter (N=", SSP_runs[[MAIN_SCENARIO]], " draws).")}
# Density plots to visualize parameter distributions
# Create facet labels
param_labels <- c("Aerosol forcing", "CO2 fertilization", "Ocean diffusivity", 
                  "ECS", "LUC emissions", "Pre-industrial NPP", "Q10")
names(param_labels) <- c("AERO_SCALE", "BETA", "DIFFUSIVITY", "ECS", "LUC_SCALE", 
                         "NPP_FLUX0", "Q10_RH")

runlist %>% 
  filter(ssp == MAIN_SCENARIO) %>%
  select(-c(run_number, ssp)) %>%
  pivot_longer(everything()) %>% 
  ggplot(aes(value)) +
  geom_density() + 
  facet_wrap(~name, scales = "free",
             labeller = labeller(name = param_labels)) +
  labs(x = NULL, y = NULL)
```


## Source coefficients of variability

``` {r s-coeff-var, eval = F, echo = FALSE, eval = FALSE, warning = FALSE, fig.cap = " Coefficients of variability for each source in the atmosphere over time."}
# Compute coefficient of variability for each source in the atmosphere pool
atmos_cv <- atmosphere_pool %>%
  filter(ssp == MAIN_SCENARIO) %>%
  group_by(year, source_name) %>%
  mutate(sdev = sd(source_fraction),
         mean = mean(source_fraction),
         cv = sdev / mean)

# Plot the coefficient of variability for each source in atmosphere pool over time
ggplot(atmos_cv, aes(year, cv, group = source_name, color = source_name)) +
  geom_line() +
  scale_color_viridis_d(breaks = c("veg_c", "detritus_c", "soil_c", "earth_c",
                                   "LL", "atmos_co2", "intermediate", "deep", "HL"),
                        labels = c("Vegetation", "Detritus", "Soil", "Anthropogenic emissions",
                                   "Low-lat ocean", "Atmosphere", "Intermed. ocean",
                                   "Deep ocean", "High-lat ocean")) +
  labs(x = "Source fraction",
       y = "Coefficient of variability",
       color = "Source parameter") +
  xlim(1800, 2300) +
  ylim(0, 0.25)

```

## Parameter distributions and source fraction controls

```{r s-parameter-space, eval = F, fig.cap = " Parameter correlations and distributions (diagonal), with color indicating how much of the atmosphere is composed of anthropogenic CO2 in the year 2100."}
# Exploring how the parameter space is linked to output
# What parameters control the percentage of earth_c in the atmosphere?
# Isolating the year 2100
pairs_labels <- c("Aerosol", "Beta", "Keff", 
                  "ECS", "LUC", "NPP", "Q10")
names(pairs_labels) <- c("AERO_SCALE", "BETA", "DIFFUSIVITY", "ECS", "LUC_SCALE", 
                         "NPP_FLUX0", "Q10_RH")

ffi_atm_2100 <- atmosphere_pool %>%
  filter(source_name == "earth_c",
         year == 2100) %>% 
  mutate(sf = cut(source_fraction, 3)) %>% 
  left_join(runlist, by = c("run_number", "ssp"))
n <- length(name_vector)
which_cols <- which(names(ffi_atm_2100) %in% names(name_vector))
ggpairs(ffi_atm_2100,
        columns = which_cols,
        # Color by magnitude of source_fraction - how much earth_c is in the atmosphere
        aes(color = sf), 
        diag = list(mapping = aes(alpha = 0.5)),
        upper = list(continuous = "blank"),
        legend = c(6,5),
        columnLabels = pairs_labels) + 
  scale_color_viridis_d() + 
  scale_fill_viridis_d() +
  labs(color = "Source Fraction")
```

## Airborne fraction window size

```{r s-af-window, eval = F, fig.cap = " Airborne fraction mathematically computed over varying time window sizes."}
# Plot airborne fraction window size plot
af_results %>% 
  group_by(year, ssp, window_size) %>% 
  summarise(af = mean(af), .groups = "drop") %>% 
  filter(year %in% 1900:2200) %>%
  ggplot(aes(year, af, color = window_size, group = window_size)) + 
  geom_line(na.rm = TRUE) + 
  facet_wrap(~ssp, scales = "free",
             labeller = labeller(ssp = ssp_labels)) +
  scale_color_viridis_c() +
  labs(x = "Year",
       y = "Airborne fraction",
       color = "Window size")

```


``` {r s-tracking-af, echo = FALSE, eval = FALSE}
# We want to calculate airborne fraction in the year immediately following the 
# tracking date. 
date_plus_one <- dates + 1
year_one <- all_years_tracking %>% 
  filter(year %in% date_plus_one) %>%
  # Assign arbitrary run number in order to use the emissions functions defined above
  mutate(run_number = 1)

# Compute cumulative emissions over time
emissions_one <- calc_emissions(year_one)

# Compute the fraction of fossil fuel emissions residing in the atmosphere
emissions_frac_one <- calc_emissions_fraction(year_one, emissions_one)

# Plot
emissions_frac_one %>%
  ggplot(aes(year, c_emissions_fraction, color = cumulative_emissions)) +
  geom_point() +
  scale_color_viridis_c()
```

# The End

```{r info}
sessionInfo()
```
