---
title: "trackingC.Rmd"
author: "Leeya Pressburger"
date: "`r Sys.Date()`"
output:
  bookdown::html_document2:
    fig_caption: yes
    number_sections: yes
    toc: true
    toc_depth: 2
    toc_float: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
```

### R setup

Load necessary packages, read in an ini file and initiate core, and set seed for reproducibility.

```{r intro, warning=FALSE, message=FALSE}
library(hector)
library(relaimpo)
library(dplyr)
library(tidyr)
library(ggplot2)
library(gridExtra)
library(ggExtra)
library(GGally)
library(bookdown)
library(data.table)

theme_set(theme_light())

source("trackingC_common.R")

# Read in data files
runlist <- fread("./output_files/runlist.csv")
model_output <- fread("./output_files/model_output.csv")
trk_output <- fread("./output_files/trk_output.csv")

# If there are over 100 observations per time point and we want to plot
# individual runs, we use slice_sample() to randomly select 100 rows.

# Pull random run numbers from the runlist and filter output by random runs
max_display <- slice_sample(runlist, n = 100)
trk_output_slice <- filter(trk_output, run_number %in% max_display$run_number)
model_output_slice <- filter(model_output, run_number %in% max_display$run_number)

```

# Results

The following sections contain tentative figures in the order they would appear
in a manuscript. 

# Primary figures

## Hector temperature vs. CMIP6 

``` {r filtered, fig.cap = "Global temperature anomaly from CMIP6 and filtered Hector runs. Lines show median; shaded areas show ±1 s.d. of median (darker) and minimum and maximum of ensemble (lighter)."}
# Read in CMIP6 comparison data
cmip_raw <- read.csv("./data_files/CMIP6_annual_tas_global.csv")

# Isolate desired scenarios
cmip_sce <- list()
for(n in names(SSP_files)){
  cmip_sce[[n]] <- cmip_raw %>% 
    filter(experiment == n)
}  
cmip_sce <- bind_rows(cmip_sce) 

# Filter for Tgav and calculate summary metrics
cmip_scenarios <- cmip_sce %>%
  rename(ssp = experiment) %>%
  filter(variable == "Tgav",
         year <= 2100) %>%
  group_by(year, ssp) %>%
  summarize(sdev = sd(value),
            minimum = min(value),
            maximum = max(value),
            med = median(value)) %>%
  mutate(source = "CMIP6") %>%
  arrange(ssp, year)

# Identify unrealistic runs - for each run, if more than 50% of values are
# outside of the min/max of the CMIP data, the run is nixed
cmip_bounds <- cmip_scenarios %>% 
  group_by(ssp) %>%
  select(year, minimum, maximum)

hector_temp <- model_output %>%
  filter(variable == "Tgav",
         year >= 2015 & year <= 2100) %>%
  select(run_number, ssp, year, value) %>%
  left_join(cmip_bounds, by = c("year", "ssp")) %>% 
  mutate(fail = value < minimum | value > maximum)

run_fail_summary <- hector_temp %>%
  group_by(ssp, run_number) %>%
  summarise(fraction_fail = sum(fail) / n())

# Filter out 'unrealistic' runs (identified above)
model_output_filtered <- model_output %>%
  left_join(run_fail_summary, by = c("ssp", "run_number")) %>% 
  filter(fraction_fail < 0.5)

# Note that the tracking output needs to be filtered as well
trk_output_filtered <- trk_output %>%
  left_join(run_fail_summary, by = c("ssp", "run_number")) %>% 
  filter(fraction_fail < 0.5)

# Create facet labels
var_labels <- c("SSP119", "SSP245", "SSP370", "SSP460", "SSP585")
names(var_labels) <- c("ssp119", "ssp245", "ssp370", "ssp460", "ssp585")

# Plot temperature spreads for the filtered runs
model_output_filtered %>%
  filter(variable == "Tgav",
         year >= 2015 & year <= 2100) %>% 
  # compute median, min, max, etc. for all remaining runs
  group_by(year, ssp) %>%
  summarize(sdev = sd(value),
            minimum = min(value),
            maximum = max(value),
            med = median(value), .groups = "drop") %>% 
  # combine with CMIP6 data
  mutate(source = "Hector") %>% 
  bind_rows(cmip_scenarios) %>%
  # ...and plot
  ggplot(aes(year, fill = source, color = source)) +
  geom_ribbon(aes(ymin = minimum, ymax = maximum), alpha = 0.15, color = NA) +
  geom_ribbon(aes(ymin = med - sdev, ymax = med + sdev), alpha = 0.5, color = NA) +
  geom_line(aes(y = med), size = 2) +
  scale_color_viridis_d(begin = 0.5) +
  scale_fill_viridis_d(begin = 0.5) +
  facet_wrap(~ssp, scales = "free",
             labeller = labeller(ssp = var_labels)) +
  #theme(legend.title = element_blank()) +
  labs(x = "Year",
       y = expression(degree*C),
       color = "Source",
       fill = "Source")
```


## Hector atmospheric CO~2~ vs observed data

``` {r ca-comp, fig.cap = " Historical CO2 data from Hector runs, NOAA observations, and CMIP6 models."}
# Comparing outputs to NOAA CO2 data from Mauna Loa observatory
# https://gml.noaa.gov/ccgg/trends/data.html
noaa <- read.csv("./data_files/noaa_co2_data.csv")
noaa_co2 <- noaa %>% 
  mutate(source = "NOAA") %>%
  filter(year < 2015)

cmip_c <- read.csv("./data_files/CMIP6_annual_co2.csv")
cmip_co2 <- cmip_c %>%
  filter(year %in% 1959:2021,
         experiment == "historical") %>%
  mutate(value = value * 1e6,
         units = "ppm") %>%
  group_by(year) %>%
  summarize(sdev = sd(value), 
            minimum = min(value),
            maximum = max(value),
            med = median(value),
            source = "CMIP6",
            .groups = "drop")

hector_co2 <- model_output_filtered %>%
  filter(variable == "Ca",
         year %in% 1959:2021) %>%
  select(run_number, year, value) %>%
  mutate(units = "ppm") %>%
  group_by(year) %>%
  summarize(sdev = sd(value), 
            minimum = min(value),
            maximum = max(value),
            med = median(value),
            source = "Hector",
            .groups = "drop")

plot_data_co2 <- bind_rows(cmip_co2, hector_co2)

plot_data_co2 %>% 
  filter(year < 2015) %>%
  ggplot(aes(year, fill = source, color = source)) +
  geom_ribbon(aes(ymin = minimum, ymax = maximum), alpha = 0.15, color = NA) +
  geom_ribbon(aes(ymin = med - sdev, ymax = med + sdev), alpha = 0.5, color = NA) +
  geom_line(aes(y = med), size = 2) +
  geom_line(data = noaa_co2, aes(year, mean, color = source), size = 2) +
  scale_color_viridis_d(begin = 0.5) +
  scale_fill_viridis_d(begin = 0.5) +
  labs(x = "Year",
       y = "CO2 (ppm)",
       color = "Source",
       fill = "Source")

```

## Atmosphere sources - ƒATM~ffi~

``` {r atmosphere-by-source, fig.cap = paste0(" Fraction of CO2 in the atmosphere by source pool from 100 random runs in ", MAIN_SCENARIO, ". Dark line shows the median.")}
# Isolate the atmosphere pool to better understand its composition 
atmosphere_pool <- trk_output_filtered %>%
  filter(pool_name == "atmos_c")

# Create facet labels
### FIX THIS
var_labels_s <- c("atmos_c", "deep", "detritus_c", "HL", "intermediate",
                               "LL", "soil_c", "veg_c")
names(var_labels_s) <- c("Atmosphere", "Deep ocean", "Detritus",
                               "High-lat ocean", "Intermed. ocean",
                               "Low-lat ocean", "Soil", "Vegetation")

# This plot looks at source fraction by source pool in the atmosphere and the median run.
atmosphere_pool %>%
  filter(ssp == MAIN_SCENARIO,
         run_number %in% max_display$run_number) %>%
  ggplot(aes(year, source_fraction, 
             color = as.factor(run_number), group = run_number)) +
  geom_line(size = 0.5, show.legend = FALSE) +
  facet_wrap(~source_name, scales = "free"
             #labeller = labeller(source_name = var_labels_s)) +
  labs(x = "Year",
       y = "Source fraction") +
  scale_color_grey(start = 0.7, end = 0.7) +
  stat_summary(color = "black",
               fun = median,
               geom = "line",
               group = "run_number",
               size = 0.7)
```

``` {r earthc-in-atmos, fig.cap = paste0(" Fraction of atmospheric CO2 from fossil fuel and industrial (FFI) emissions over time in ", MAIN_SCENARIO, ". Line shows median; shaded areas show ±1 s.d. of median (darker) and minimum and maximum of ensemble (lighter).")}
# This chunk looks at the atmosphere pool with just one source (earth_c) 
# and plots source fraction over time with a confidence interval

ff <- atmosphere_pool %>% 
  filter(source_name == "earth_c") %>%
  group_by(ssp, year) %>%
  mutate(sf_sd = sd(source_fraction), 
            sf_min = min(source_fraction),
            sf_max = max(source_fraction),
            sf_median = median(source_fraction))

ggplot(ff, aes(year, color = ssp, fill = ssp)) +
  geom_line(aes(y = sf_median), size = 2) +
  geom_ribbon(aes(ymin = sf_min, ymax = sf_max), alpha = 0.2) +
  geom_ribbon(aes(ymin = sf_median - sf_sd, ymax = sf_median + sf_sd), alpha = 0.2) +
  scale_color_viridis_d(breaks = c("ssp119", "ssp245", "ssp370", "ssp460", "ssp585"),
                        labels = c("SSP119", "SSP245", "SSP370", "SSP460", "SSP585")) + 
  scale_fill_viridis_d(breaks = c("ssp119", "ssp245", "ssp370", "ssp460", "ssp585"),
                        labels = c("SSP119", "SSP245", "SSP370", "SSP460", "SSP585")) +
  labs(x = "Year",
       y = "Fraction of atmospheric CO2",
       color = "Scenario",
       fill = "Scenario")
```

## Destination of FFI emissions - ƒFFI~atm~

```{r destination, fig.cap = " Final destination pools of fossil fuel and industrial (FFI) emissions for 16 random runs."}
# Where does CO2 end up?
# Step 1: compute source_quantity = pool_value * source_fraction
# Step 2: for each year and source_name, compute destination_fraction

ffi_destinations <- trk_output_filtered %>%
  filter(source_name == "earth_c", pool_name != "earth_c") %>%
  mutate(source_quantity = pool_value * source_fraction) %>%
  group_by(year, ssp, run_number) %>%
  mutate(destination_fraction = source_quantity / sum(source_quantity)) %>% 
  ungroup()

# Just looking at 5 random runs across scenarios
### THIS NO LONGER WORKS
dest_runs <- ffi_destinations %>%
  group_by(ssp) %>%
  filter(year == min(year)) %>%
  slice_sample(n = 5) %>%
  summarize(run_number)

dest_plot <- filter(ffi_destinations, 
                    run_number %in% dest_runs$run_number) 

# Just looking at the mean of all runs in SSP245
dest_plot %>%
  filter(ssp == MAIN_SCENARIO) %>%
  group_by(ssp, year, pool_name) %>%
  mutate(dest_avg = mean(destination_fraction)) %>%
  filter(run_number == min(run_number)) %>%
  ggplot(aes(year, dest_avg, fill = pool_name)) +
  geom_area() +
  scale_fill_viridis_d(breaks = c("atmos_c", "deep", "detritus_c", "HL", "intermediate",
                               "LL", "soil_c", "veg_c"),
                    labels = c("Atmosphere", "Deep ocean", "Detritus",
                               "High-lat ocean", "Intermed. ocean",
                               "Low-lat ocean", "Soil", "Vegetation")) +
  # facet_wrap(~ssp) +
  #theme(#strip.background = element_blank(),
        #strip.text.x = element_blank(),
        #axis.text.x = element_text(angle = 90)) +
  labs(x = "Year",
       y = "Fraction of FFI emissions",
       fill = "Destination")
```

```{r relative-importance, message=FALSE, echo=FALSE, warning=FALSE, fig.cap = " Relative importance of parameters over time in controlling atmosphere as a destination of fossil fuel and industrial (FFI) emissions."}
# Filter to just atmosphere pool, and remove unneeded columns
dest_e2a <- filter(ffi_destinations, pool_name == "atmos_c")
dest_e2a_minimal <- dest_e2a[c("year", "destination_fraction", "ssp", names(name_vector))]

# Function to calculate relative importance of parameters in
# explaining how much destination_fraction varies for a given
# year of data. This uses relaimpo::calc.relimp()
# See https://www.r-bloggers.com/2012/08/the-relative-importance-of-predictors-let-the-games-begin/
# for a useful primer, along with the calc.reimp help page
calc_relimp <- function(x) {
  yr <- x$year[1] # record year...
  x$year <- NULL # and drop year column
  sce <- x$ssp[1]
  x$ssp <- NULL
  # Fit model. The formula notation means "destination fraction as a
  # function of all other terms" (which is why we dropped year)
  m <- lm(destination_fraction ~ ., data = x)
  # Calculate relative importance metrics and extract 'lmg' results
  # 'lmg' is the R^2 contribution averaged over orderings among regressors;
  # should sum to one because we're using relative importances (rela=TRUE)
  relimp <- try(calc.relimp(m, type = "lmg", rela = TRUE)@lmg)
  if(class(relimp) == "try-error") {
    message(yr, " calc.relimp error")
    return(NULL)
  }
  # Return a data frame: year, parameter, relative importance
  tibble(ssp = sce, year = yr, param = names(relimp), value = relimp)
  
}

# Run calc_relimp() above on each year's data
dest_split <- split(dest_e2a_minimal, list(dest_e2a_minimal$year, dest_e2a_minimal$ssp))
relimp_out <- lapply(dest_split, FUN = calc_relimp) %>% 
  bind_rows()
    
# ...and plot!
ggplot(relimp_out, aes(year, value, fill = param)) + 
  geom_area() +
  coord_cartesian(expand = FALSE) +
  scale_fill_viridis_d(direction = -1,
                       breaks = c("AERO_SCALE", "BETA", "DIFFUSIVITY",
                                  "ECS", "LUC_SCALE", "NPP_FLUX0", "Q10_RH"),
                       labels = c("Aerosol forcing", "CO2 fertilization",
                                  "Ocean diffusivity", "ECS",
                                  "LUC emissions", "Pre-industrial NPP",
                                  "Q10")) +
  facet_wrap(~ssp,
             labeller = labeller(ssp = var_labels)) +
  labs(x = "Year",
       y = expression(Relative~importance~"for"~FFI[atm]),
       fill = "Parameter") +
  theme(axis.text.x = element_text(angle = 90))
```

## Does time affect destination?

```{r tracking-multiple-years, warning = FALSE, fig.cap = " Amount of CO2 in the atmosphere sourced from human emissions. Note that turning on tracking in different years does not have a dramatic effect on the composition of the atmosphere pool."}
# Prepare data: reset the core, set start dates every 50 years
tcore <- newcore(SSP_files[[MAIN_SCENARIO]])
dates <- seq(from = 1750, to = 2250, 50)
tracking_output <- list()

# Function to set a tracking date, run the core for a 50 year period, and retrieve tracking data
set_tracking <- function(c, date){
  setvar(c, NA, TRACKING_DATE(), date, "(unitless)")
  reset(c)
  run(c, (date + 50))
  tdata <- get_tracking_data(c)
  tdata$ty <- date
  tdata$ssp <- MAIN_SCENARIO
  tdata
}

# Loop through above function for desired dates
for(d in dates){
  tracking_output[[d]] <- set_tracking(tcore, d)
}

all_years_tracking <- bind_rows(tracking_output)

# Plot the fraction of the atmosphere that is sourced from earth_c
# all_years_tracking %>%
#   filter(source_name == "earth_c",
#          pool_name == "atmos_c") %>%
#   ggplot(aes(year, source_fraction, color = ty, group = ty)) +
#   geom_line() +
#   labs(x = "Year",
#        y = "Source fraction") +
#   scale_color_viridis_c()

# Plot the destination of earth_c emissions
all_years_tracking %>%
  filter(source_name == "earth_c", pool_name != "earth_c") %>%
  mutate(source_quantity = pool_value * source_fraction) %>%
  group_by(year) %>%
  mutate(destination_fraction = source_quantity / sum(source_quantity)) %>% 
  ungroup() %>%
  ggplot(aes(year, destination_fraction, fill = pool_name), group = ty) +
  geom_area() +
  labs(x = "Year",
       y = "Destination fraction",
       fill = "Parameter") +
  scale_fill_viridis_d(breaks = c("atmos_c", "deep", "detritus_c", "HL", "intermediate",
                               "LL", "soil_c", "veg_c"),
                    labels = c("Atmosphere", "Deep ocean", "Detritus",
                               "High-lat ocean", "Intermed. ocean",
                               "Low-lat ocean", "Soil", "Vegetation"))

```

## Airborne fraction

```{r airborne, warning = FALSE, fig.cap = " Year-by-year mathematically defined airborne fraction. The yellow line and band shows the actual amount of human emissions remaining in the atmosphere from the model's carbon-tracking mechanism."}
# Exploring how the parameter space is linked to output
# What parameters control the percentage of earth_c in the atmosphere?
# Isolating the year 2100

ffi_atm_2100 <- atmosphere_pool %>%
  filter(source_name == "earth_c",
         year == 2100) %>% 
  mutate(sf = cut(source_fraction, 3))

n <- length(name_vector)
ggpairs(ffi_atm_2100,
        columns = 10:(10 + (n-1)),
        # Color by magnitude of source_fraction - how much earth_c is in the atmosphere
        aes(color = sf), 
        diag = list(mapping = aes(alpha = 0.5)),
        upper = list(continuous = "blank"),
        legend = c(6,5)) + 
  scale_color_viridis_d() + 
  scale_fill_viridis_d() +
  labs(color = "Source Fraction")
```

## Airborne fraction time windows

```{r af-windows, fig.cap = " Airborne fraction computed over varying time windows."}
# Prep the data
atm_earth_diffs <- trk_output_filtered %>%
  # Isolate atmosphere and earth total pool values
  filter(pool_name %in% c("atmos_c", "earth_c")) %>% 
  group_by(run_number, year, pool_name, ssp) %>% 
  summarise(pool_value = mean(pool_value), .groups = "drop") %>% 
  arrange(year) %>%
  # Put atmosphere and earth in separate columns and compute diffs
  pivot_wider(names_from = "pool_name", values_from = "pool_value") %>% 
  group_by(run_number, ssp) %>%
  mutate(atm_diff = c(NA, diff(atmos_c)), 
         earth_diff = c(NA, diff(earth_c)))

# Compute AF (classical definition of delta CO2atm / emissions) 
# using various window sizes
af_results <- list()
for(groupsize in seq(1, 40, length.out = 10)) {
  ngroups = length(unique(atm_earth_diffs$year)) / groupsize
  
  atm_earth_diffs %>% 
    group_by(run_number, ssp) %>% 
    mutate(group = cut(year, breaks = ngroups)) %>% 
    # compute differences over the group of years
    group_by(run_number, group, ssp) %>%
    summarise(year = max(year), 
              atm_diff = sum(atm_diff), 
              earth_diff = sum(earth_diff), 
              n = n(),
              af = atm_diff / -earth_diff,
              .groups = "drop") ->
    af_results[[as.character(groupsize)]]
}

af_results <- bind_rows(af_results, .id = "window_size") %>% 
  mutate(window_size = as.integer(window_size))


# Compute cumulative emissions over time
# We use the change in earth_c to compute emissions; note this will not work
# if ever using a scenario with fossil fuel sequestration (atmosphere -> earth)

# This function computes annual cumulative earth_c emissions for a given tracking dataset
min_runs <- trk_output_filtered %>%
  group_by(ssp) %>%
  summarize(mrun = min(run_number))

calc_emissions <- function(data){
  data %>% 
    filter(pool_name == "earth_c",
         source_name == "earth_c",
         run_number %in% min_runs$mrun) %>%
    arrange(year) %>%
    group_by(ssp) %>%
    # the NA at the end here is so that cumulative emissions 'line up' correctly
    # with the source_quantity (earth-origin atmosphere C) calculated below
    mutate(emissions = -c(diff(pool_value), NA),
           # relative to core$trackingDate
           cumulative_emissions = cumsum(emissions)) %>%
    select(year, ssp, cumulative_emissions) %>%
    arrange(ssp, year)
}


# This function takes a given tracking dataset and its corresponding cumulative 
# emissions dataset as found above to calculate what fraction of the cumulative
# emissions come from earth_c emisions that are now in the atmosphere
calc_emissions_fraction <- function(data, emissions_data){
  data %>%
    filter(pool_name == "atmos_c", 
           source_name == "earth_c") %>%
    # How much earth_c is in atmos_c? 
    mutate(source_quantity = pool_value * source_fraction) %>%
    left_join(emissions_data, by = c("year", "ssp")) %>%
    select(run_number, ssp, year, source_quantity, cumulative_emissions) %>%
    mutate(c_emissions_fraction = source_quantity / cumulative_emissions) %>% 
    group_by(year, ssp) %>%
    mutate(sdev = sd(c_emissions_fraction),
           med = median(c_emissions_fraction),
           minimum = min(c_emissions_fraction),
           maximum = max(c_emissions_fraction),
           def = "EmissFraction")
}

# Compute cumulative emissions over time
emissions <- trk_output_filtered %>% calc_emissions() 

# Compute the fraction of fossil fuel emissions residing in the atmosphere
emissions_fraction <- calc_emissions_fraction(trk_output_filtered, emissions)

# Classical definition of AF - find summary stats
classic_af <- af_results %>%
  group_by(group, year, ssp) %>%
  mutate(sdev = sd(af, na.rm = TRUE),
         minimum = min(af),
         maximum = max(af),
         med = median(af),
         def = "ClassicAF") %>% 
  filter(window_size == 1,
         run_number %in% emissions_fraction$run_number)

# Isolate relevant parameters and plot
plot_data <- bind_rows(emissions_fraction, classic_af) %>%
  select(run_number, year, ssp, minimum, maximum, med, sdev, def) %>%
  filter(year %in% 1910:2200) # past 2200, data gets weird

ggplot(plot_data, aes(x = year, fill = def)) +
  geom_ribbon(aes(ymin = minimum, ymax = maximum), 
              alpha = 0.2) +
  geom_ribbon(aes(ymin = med - sdev, ymax = med + sdev), 
              alpha = 0.4) +
  geom_line(aes(y = med, color = def)) +
  scale_color_viridis_d(begin = 0.5,
                        breaks = c("ClassicAF", "EmissFraction"),
                        labels = c("Traditional",
                                   "Carbon tracking")) +
  scale_fill_viridis_d(begin = 0.5,
                        breaks = c("ClassicAF", "EmissFraction"),
                        labels = c("Traditional",
                                   "Carbon tracking")) +
  facet_wrap(~ssp, scales = "free_y", labeller = labeller(ssp = var_labels)) +
  labs(x = "Year",
      y = "Airborne fraction",
      color = "Computation",
      fill = "Computation") +
  theme(axis.text.x = element_text(angle = 90))

```


# Supplementary figures

## Realized parameter PDFs

``` {r params-PDFs, fig.cap = paste0(" Probability densities of each parameter (N=", SSP_runs[[MAIN_SCENARIO]], " draws).")}
# Density plots to visualize parameter distributions
runlist %>% 
  filter(ssp == MAIN_SCENARIO) %>%
  select(-c(run_number, ssp)) %>%
  pivot_longer(everything()) %>% 
  ggplot(aes(value)) +
  geom_density() + 
  facet_wrap(~name, scales = "free") +
  labs(x = NULL, y = NULL)
```

## Hector temperature vs CMIP6 diagnostic

``` {r cmip, message = FALSE, fig.cap = " Visualization of sample model runs. The black dots represent the CMIP6 minimums and maximums for global temperature anomaly, and the runs that pass must fall within these bounds more than 50% of the time."}
# Visualize the run fail summary calculated above
# Make a table of failure rates
run_fail_summary %>% 
  group_by(fraction_fail > 0.5) %>% 
  summarise(n()) %>% 
  knitr::kable()

# ...and a figure 
hector_temp_runs <- hector_temp %>% 
  group_by(ssp) %>%
  filter(year == min(year)) %>%
  slice_sample(n = 100) %>% # this is a debugging figure; plot just 100 runs for clarity
  arrange(ssp, run_number) %>%
  summarize(run_number)

# Filter out data for just the 100 rusn per senario identified above
hector_temp_slice <- left_join(hector_temp_runs, hector_temp, by = c("run_number", "ssp"))
  
# Plot
left_join(hector_temp_slice, run_fail_summary, by = c("run_number", "ssp")) %>%
  ggplot(aes(year, value, color = fraction_fail, group = run_number)) + 
  geom_line() + 
  geom_point(aes(y = minimum), color = "black") + 
  geom_point(aes(y = maximum), color = "black") +
  facet_wrap(~ssp,
             labeller = labeller(ssp = var_labels)) +
  scale_color_viridis_c() +
  labs(x = "Year",
       y = expression(degree*C),
       color = "Fraction of failed runs")

```

## Source coefficients of variability

``` {r coeff-var, fig.cap = " Coefficients of variability for each source in the atmosphere over time."}
# Compute coefficient of variability for each source in the atmosphere pool
atmos_cv <- atmosphere_pool %>%
  filter(ssp == MAIN_SCENARIO) %>%
  group_by(year, source_name) %>%
  mutate(sdev = sd(source_fraction),
         mean = mean(source_fraction),
         cv = sdev / mean)

# Plot the coefficient of variability for each source in atmosphere pool over time
ggplot(atmos_cv, aes(year, cv, group = source_name, color = source_name)) +
  geom_line() +
  scale_color_viridis_d(breaks = c("atmos_c", "deep", "detritus_c", "HL", "intermediate",
                               "LL", "soil_c", "veg_c"),
                    labels = c("Atmosphere", "Deep ocean", "Detritus",
                               "High-lat ocean", "Intermed. ocean",
                               "Low-lat ocean", "Soil", "Vegetation")) +
  labs(x = "Source fraction",
       y = "Coefficient of variability",
       color = "Source parameter") +
  xlim(1800, 2300) +
  ylim(0, 0.25)

```

## Parameter distributions and source fraction controls

```{r parameter-space, fig.cap = " Parameter correlations and distributions (diagonal), with color indicating how much of the atmosphere is composed of anthropogenic CO2 in the year 2100."}
# Exploring how the parameter space is linked to output
# What parameters control the percentage of earth_c in the atmosphere?
# Isolating the year 2100
ffi_atm_2100 <- atmosphere_pool %>%
  filter(source_name == "earth_c",
         year == 2100) %>% 
  mutate(sf = cut(source_fraction, 3))
n <- length(name_vector)
ggpairs(ffi_atm_2100,
        columns = 10:(10 + (n-1)),
        # Color by magnitude of source_fraction - how much earth_c is in the atmosphere
        aes(color = sf), 
        diag = list(mapping = aes(alpha = 0.5)),
        upper = list(continuous = "blank"),
        legend = c(6,5)) + 
  scale_color_viridis_d() + 
  scale_fill_viridis_d() +
  labs(color = "Source Fraction")
```

# Airborne fraction window size

```{r parameter-space, fig.cap = " Parameter correlations and distributions (diagonal), with color indicating how much of the atmosphere is composed of anthropogenic CO2 in the year 2100."}
# Plot airborne fraction window size plot
af_results %>% 
  group_by(year, ssp, window_size) %>% 
  summarise(af = mean(af), .groups = "drop") %>% 
  filter(year %in% 1900:2200) %>%
  ggplot(aes(year, af, color = window_size, group = window_size)) + 
  geom_line(na.rm = TRUE) + 
  facet_wrap(~ssp, scales = "free",
             labeller = labeller(ssp = var_labels)) +
  scale_color_viridis_c() +
  labs(x = "Year",
       y = "Airborne fraction",
       color = "Window size") 

```

## Tracking year plus one (work in progress)

``` {r tracking-af}
# We want to calculate airborne fraction in the year immediately following the 
# tracking date. 
date_plus_one <- dates + 1
year_one <- all_years_tracking %>% 
  filter(year %in% date_plus_one) %>%
  # Assign arbitrary run number in order to use the emissions functions defined above
  mutate(run_number = 1)

# Compute cumulative emissions over time
emissions_one <- calc_emissions(year_one)

# Compute the fraction of fossil fuel emissions residing in the atmosphere
emissions_frac_one <- calc_emissions_fraction(year_one, emissions_one)

# Plot
emissions_frac_one %>%
  ggplot(aes(year, c_emissions_fraction, color = cumulative_emissions)) +
  geom_point() +
  scale_color_viridis_c()
```

# The End

```{r info}
sessionInfo()
```
